Efficient and Accurate Modeling of Flow through Sedimentary Structures with MODFLOW 6

### 2023-10-19 latex/article.tex
abstract and article impact statement

\textbf{Article impact statement:} {\color{red} AMP: Need impact statement}

\begin{abstract}
\noindent {\color{red} AMP: compose Abstract near the end}
\end{abstract}

>>>>

\textbf{Article impact statement: Groundwater flow through dipping sedimentary structures can be accurately simulated using the XT3D multi-point flow formulation in MODFLOW 6.}

\begin{abstract}
Groundwater flow through dipping hydrogeologic layers can be modeled with MODFLOW by specifying top and bottom elevations of model layers to coincide with top and bottom elevations of hydrogeologic layers.  With this approach, adjacent cells within a model layer are vertically offset from one another.  While the standard formulation in MODFLOW does not strictly handle these offsets in the underlying flow calculations, results are generally considered to be accurate provided dip angles do not exceed $10^{\circ}$.  To improve the accuracy of flow simulations on unstructured grids, and to allow accurate simulation of arbitrarily oriented two- or three-dimensional anisotropy, the optional XT3D multi-point flow formulation was introduced in MODFLOW 6.  The XT3D flow formulation was designed to handle grid irregularities, such as vertical offsets between laterally adjacent cells, and to provide accurate results for both isotropic and anisotropic groundwater flow.  \cite{bardot2022} showed, however, that XT3D simulations did not produce the correct flow direction and magnitude for simple tests involving flow through a steeply dipping permeable channel.  In this paper, we identify the cause of the poor XT3D performance reported by \cite{bardot2022} for the steeply dipping test problem: for XT3D to perform as intended, the full cell connectivity is required for vertically offset grids.  The common discretization packages in MODFLOW 6 are based on layered connectivity in which cells are laterally connected only with cells in the same model layer.   When the fully unstructured grid type in MODFLOW 6 is used for this problem, and when additional cross connections are specified to represent lateral connections between cells in different model layers, XT3D is shown to give the correct solution both in terms of flow magnitude and direction.  An important requirement for this improved accuracy, which is available for both isotropic and anisotropic conditions, is that permeable hydrogeologic features must be represented with at least two model layers.  The combined use of vertically offset grids and the XT3D formulation in MODFLOW 6 provides a computationally efficient and accurate method for simulating isotropic and anisotropic flow through sedimentary structures, a capability that was previously reserved to finite-element simulators.
\end{abstract}
